---
title: "D474_DataClean"
author: "Alex Marcek"
output: html_document
---

## Import Data

DISCLAIMER: The following chunk is included because of the structure of our directory. We want directory paths to be resolved from our project root. This does not always happen automatically depending on how our working directory is set up. A common (local) fix is found in your RStudio's options. Tools -> Global Options -> RMarkdown -> Evaluate Chunks in Directory : Project (default is Document, switch to Project). Hopefully this is not required, but it is the fix I recommend if you have trouble importing our data. 

The fix that will hopefully work for all of us is in the package `here`. Calling `here()` resolves the path relative to our project root for all repo users. Please contact Alex if importing issues persist past this documentation!

```{r}
# Force dependency
if (!require(here)) install.packages("here")
```

```{r}
# Relative path approach
library(here)
clickstream <- read.table(here("Data", "RawData", "e-shop_clothing2008.csv"),
                             header = TRUE,
                             sep = ";")

clickstream[1:5, ]

# If this chunk returns error, try setting below chunk (eval = TRUE), then run. 
```

```{r, eval = FALSE}
# RStudio path relative to project root
## Working directory error diagnosis: print any errors you receive
print(getwd())
print(file.exists("Data/RawData/e-shop_clothing2008.csv"))
clickstream <- read.csv("Data/RawData/e-shop_clothing2008.csv",
                        header = TRUE, 
                        sep = ";")
```
Our data was sourced from kaggle: https://www.kaggle.com/datasets/tunguz/clickstream-data-for-online-shopping?resource=download
This data set contains information from an online store that offers clothing for pregnant women. The data is from a five month period in 2008. The data set contains relevant (front end) information on the website, along with product information and user information. Our aim is to show how website layout and product attributes may impact user interactions . 

\newpage

## Data Overview

We will begin by examining and commenting on the structure of our data set
```{r}
# Structure
str(clickstream)
```
As we can see, the data set is a 165474 x 14 matrix. There are 14 variables with 165474 observations. 

An important note on data types. All of our variables (except for "page.2..clothing.model.", which is a character type) are integers. However, we cannot treat them all as integers. Looking at the data set's documentation, the following variables are encoded categorical. (Algebraic manipulation will be invalid for the following): `order(?)`, `country`, `session.ID`, `page.1..main.category.`, `page.2..clothing.model.`, `colour`, `location`, `model.photography`, `price.2`. So, it really only makes sense to potentially perform such algebraic manipulations on the following variables: `year`, `month`, `day`, `price`, `page`.

We will take a quick look at a random sample of our data to see if there are any surface inconsistencies in our data set. Specifically, we are looking for any missing data entries, mixed types, broken characters, implicit nulls that are not flagged as NA, and any structural surprises that may lend additional context to the way our data set functions. 
```{r}
# Random Missingness/Invalidity Audit

library(dplyr)
set.seed(050326)

# First 10 rows
head(clickstream, 10)

# 50 randomly sampled points from data 
clickstream %>% slice_sample(n = 50)

# Last 10 Rows
tail(clickstream, 10)
```

Next, we will check the ranges, or levels (for factors) of our data set. We will cross check with the sampled data above to make sure there are not any values that fall outside of feasible values. 

```{r}
summary(clickstream)
```
For ease of manipulation later on, lets take a moment to standardize the variable names: 
```{r}
name_map <- c(sid = "session.ID", main_category = "page.1..main.category.", clothing_model = "page.2..clothing.model.", color = "colour", model_photography = "model.photography", price_2 = "price.2")

clickstream <- clickstream |>
  rename(any_of(name_map))

colnames(clickstream)
```

Now, lets take note of the levels of each factor we have in this data set. We may want to convert some of these variables to factors later in our data analysis as well. For this reason, we will create a conversion function that we can utilize later in our analysis.
```{r}
col_to_factor <- c("country")
```

## Data Quality Assessment