---
title: "Exploratory Data Analysis"
Author: "Alex Marcek"
output: html_document
---

# Data Import
We will be using the cleaned data that resulted from the processing we did in `D474_DataClean.rmd`

```{r}
library(here)

clickstream <- read.csv(here("Data", "EditedData", "clickstream_dc.csv"))

head(clickstream)
str(clickstream)
```
Note that data has been interpreted back into a character data type. We will need to convert back to date if we wish to reference it as such. (We will need this later in the document).

```{r}
clickstream$date <- as.Date(clickstream$date)
```

# Univariate Analysis

```{r}
par(mfrow = c(1, 2))

barplot(
  table(as.integer(format(clickstream$date, "%m"))),
  main = "Frequency of price_2",
  xlab = "Category",
  ylab = "Count",
  col = "lightblue"
)

barplot(
  table(clickstream$country),
  main = "Frequency of price_2",
  xlab = "IP",
  ylab = "Count",
  col = "lightblue"
)
as.data.frame(table(clickstream$country)) %>% arrange(desc(Freq)) %>% head(5)
```
Months have roughly the same number of interactions.
Top 5 most popular IPs go to: Poland, Czech Republic, Lithuania, NA, NA.

# Multivariate Analysis