---
title: "Logistic Regression"
author: "Graham I. Gordon"
date: "2026-05-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(boot)
library(tidyverse)
library(here)
library(caret)
library(BART)
library(randomForest)
library(pls)
library(pROC)

library(stringr)
library(ggplot2)

library(tidymodels)
library(embed)
library(xgboost)
library(vip)

library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```


```{r}
#clickstream <- read.csv(here("Data", "RawData", "e-shop_clothing2008.csv"),
#                        header = TRUE,
#                        sep = ";")



clickstream$price_2 <- as.factor( ifelse(clickstream$price_2 == 2, 0, clickstream$price_2) )

clickstream <- clickstream %>% mutate(across(where(is.character), as.factor)) 

clickstream$price <- log(clickstream$price)
clickstream$month <- factor( format(as.Date(clickstream$date), '%m' ) )

colnames(clickstream)
```

```{r}
train <- sample( 1:nrow(clickstream), size = .7*nrow(clickstream) )
train_df <- clickstream[train, ]
test_df <- clickstream[-train, ]
table(clickstream$price_2)
prop.table(table(clickstream$hi_clicks_avg))
prop.table(table(clickstream$hi_clicks_med))
```


```{r}
logmod <- glm( price_2 ~ main_category + as.factor(color) + location, family = binomial, data = clickstream )
#cv.log <- cv.glm( clickstream, logmod, K = 6 )

logprob <- predict(logmod, test_df, type = "response")
logpred <- ifelse(logprob > 0.5, 1, 0)

confusionMatrix(factor(logpred),
                factor(test_df$price_2))
```

```{r}
summary(logmod)
```


```{r}
factor_cols <- c("main_category", "color", "location", "country", "page", "month")

# Convert in train and test
train_df[, factor_cols] <- lapply(train_df[, factor_cols], factor)
test_df[, factor_cols]  <- lapply(test_df[, factor_cols], factor)

# Then run your existing code as-is
#keep_cols <- c("main_category", "color", "location", "price", 
#               "model_photography", "page", "month")
keep_cols <- c("main_category", "color", "location", "country", "page", "month")
xtrain <- train_df[, keep_cols]
ytrain <- train_df[, "hi_clicks_med"] %>% factor()
xtest  <- test_df[, keep_cols]
ytest  <- test_df[, "hi_clicks_med"] %>% factor()

```

```{r}
# Fix levels - test adopts train's levels; test-only levels become NA
for (col in factor_cols) {
  test_df[[col]] <- factor(test_df[[col]], levels = levels(train_df[[col]]))
}

# Rebuild after fixing
xtrain <- train_df[, keep_cols]
xtest  <- test_df[, keep_cols]
ytrain <- train_df[["hi_clicks_med"]] %>% factor()
ytest  <- test_df[["hi_clicks_med"]]  %>% factor()
```


```{r}
#barfit <- pbart( xtrain,as.numeric(as.character(ytrain)), x.test = xtest, 
#                 k = 2, ntree = 200)
barfit <- pbart( xtrain,as.numeric(as.character(ytrain)), x.test = xtest, 
                 k = 2, ntree = 200, sparse = TRUE, nskip  = 200, ndpost = 1300 )
```
```{r}
library(tidyr)

# Sample 6 random test observations
set.seed(42)
sample_obs <- sample(ncol(barfit$yhat.test), 6)

pnorm(barfit$yhat.test[, sample_obs]) %>%
  as.data.frame() %>%
  setNames(paste0("Obs ", sample_obs)) %>%
  mutate(iteration = row_number()) %>%
  pivot_longer(-iteration, names_to = "observation", values_to = "prob") %>%
  ggplot(aes(x = iteration, y = prob, color = observation)) +
  geom_line(linewidth = 0.4, alpha = 0.7) +
  facet_wrap(~ observation, ncol = 2) +
  labs(
    title = "MCMC Trace Plots — Posterior Predicted Probabilities",
    x     = "MCMC Iteration",
    y     = "P(hi_clicks_avg = 1)"
  ) +
  theme_bw() +
  theme(legend.position = "none")
```

```{r}
formals(pbart)
#formals(pbart)$k      # → 2   (leaf shrinkage)
#formals(pbart)$power  # → 2   (tree depth decay)
#formals(pbart)$base   # → 0.95
#formals(pbart)$ntree  # → 200
```





```{r}
# Get predicted probabilities for test set
bart_probs <- barfit$prob.test.mean

# Convert to binary predictions
bart_pred <- ifelse(bart_probs > 0.5, 1, 0)

# Confusion matrix
confusionMatrix(factor(bart_pred), factor(ytest), positive = '1')


roc_bart <- roc(as.numeric(as.character(ytest)), bart_probs)
auc(roc_bart)
plot(roc_bart, main = paste0("BART ROC Curve | AUC = ", round(auc(roc_bart),3)))
```

For median: Accuracy : 0.6257  
For average: Accuracy : 0.6945 

Without sparse, AUC = 0.6583
With sparse, AUC =  0.6589


```{r}
# 1. Build xtrain/xtest AFTER all preprocessing
keep_cols <- c("main_category", "color", "location", "country", "page", "month")

xtrain <- data.frame(lapply(train_df[, keep_cols], as.numeric))
xtest  <- data.frame(lapply(test_df[, keep_cols],  as.numeric))
ytrain <- factor(train_df$hi_clicks_avg)
ytest  <- factor(test_df$hi_clicks_avg)

# 2. Verify dimensions before fitting
dim(xtrain)   # e.g. 104248 x 7
dim(xtest)
levels(ytrain)  # "0" "1"

# 3. Fit
plsda.fit <- train(x          = xtrain,
                   y          = ytrain,
                   method     = "pls",
                   tuneLength = 5,
                   trControl  = trainControl(method = "cv", number = 10),
                   preProcess = c("center", "scale"))

# 4. Predict and evaluate
pred_class <- predict(plsda.fit, newdata = xtest)
confusionMatrix(pred_class, ytest)
```

```{r}


# Your known original variable names
original_vars <- c("main_category", "color", "location", "country", "page", "month")

# Extract and aggregate importance
importance_df <- data.frame(
  dummy_var  = colnames(barfit$varcount),
  importance = colMeans(barfit$varcount)
) %>%
  mutate(original_var = case_when(
    dummy_var %in% original_vars ~ dummy_var,
    TRUE ~ str_remove(dummy_var, "\\d+$")
  )) %>%
  group_by(original_var) %>%
  summarise(importance = sum(importance)) %>%
  mutate(importance_pct = importance / sum(importance) * 100) %>%
  arrange(desc(importance))

print(importance_df)

ggplot(importance_df, aes(x = reorder(original_var, importance), y = importance_pct)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "BART Variable Importance (aggregated)",
    x = NULL,
    y = "% of total splits"
  ) +
  theme_minimal()
```



```{r}
# Step 1: normalize each posterior draw to proportions (ndpost x 37)
var_props <- t(apply(barfit$varcount, 1, function(x) x / sum(x)))

# Step 2: convert to data frame with dummy var names
var_props_df <- as.data.frame(var_props)
colnames(var_props_df) <- colnames(barfit$varcount)

# Step 3: aggregate dummy columns back to original variables for each posterior draw
# Transpose so vars are rows, then group
agg_props <- var_props_df %>%
  t() %>%
  as.data.frame() %>%
  mutate(dummy_var = colnames(barfit$varcount)) %>%
  mutate(original_var = case_when(
    dummy_var %in% original_vars ~ dummy_var,
    TRUE ~ str_remove(dummy_var, "\\d+$")
  )) %>%
  group_by(original_var) %>%
  summarise(across(where(is.numeric), sum))

# Step 4: compute posterior mean and credible intervals across draws
draws_only <- agg_props %>% select(-original_var) %>% as.matrix()

var_df <- data.frame(
  var   = agg_props$original_var,
  mean  = rowMeans(draws_only),
  lower = apply(draws_only, 1, quantile, 0.05),
  upper = apply(draws_only, 1, quantile, 0.95)
)

# Step 5: plot
ggplot(var_df, aes(x = mean, y = reorder(var, mean))) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) +
  labs(title = "Variable Importance with 90% Credible Intervals",
       x = "Posterior split proportion", y = NULL) +
  theme_minimal()
```


Random Forest:
```{r}
rf_fit <- randomForest(x = xtrain, y = factor(ytrain),
                       ntree = 500,
                       sampsize = c("0" = 1000, "1" = 1000),
                       importance = TRUE)

# Get predicted probabilities for class 1
rf_prob <- predict(rf_fit, xtest, type = "prob")[, 2]

# ROC curve
roc_rf <- roc(as.numeric(as.character(ytest)), rf_prob)

# Plot it
plot(roc_rf, main = paste("Random Forest ROC — AUC:", round(auc(roc_rf), 3)))

bart_pred <- ifelse(bart_probs > 0.5, 1, 0)

confusionMatrix(factor(bart_pred), factor(ytest))
```
Random forest Accuracy : 0.6497 

Boosting:
```{r}
library(xgboost)

xgb_fit <- xgboost(
  x = as.matrix(xtrain %>% mutate(across(where(is.factor), as.numeric))),
  y = factor(as.numeric(as.character(ytrain))),
  nrounds = 100,
  objective = "binary:logistic",
  scale_pos_weight = sum(ytrain == 0) / sum(ytrain == 1),
  learning_rate = 0.1
)
```

```{r}
# Get predicted probabilities
xgb_prob <- predict(xgb_fit, 
                    newdata = as.matrix(xtest %>% mutate(across(where(is.factor), as.numeric))))

# Convert to binary predictions
xgb_pred <- ifelse(xgb_prob > 0.5, 1, 0)

# Confusion matrix
confusionMatrix(factor(xgb_pred), factor(as.numeric(as.character(ytest))))

# AUC
roc_xgb <- roc(as.numeric(as.character(ytest)), xgb_prob)
auc(roc_xgb)
plot(roc_xgb, main = paste("XGBoost ROC — AUC:", round(auc(roc_xgb), 3)))
```

Gradient Boosting Accuracy : 0.6180 





Switch to target encoding to include clothing_model

Here is a gradient boosted tree ensemble model which uses target encoding so we can include clothing_model without needing 52 extra dummy variables.

```{r}
library(tidymodels)
library(embed)
library(xgboost)
library(vip)

# ── 1. Prep ───────────────────────────────────────────────────────────────────

clickstream_model <- clickstream %>%
  mutate(
    hi_clicks_avg = as.factor(hi_clicks_avg),  # must be factor for classification
    across(c(main_category, color, location, model_photography, 
             page, month, clothing_model), as.factor)
  )

set.seed(42)
split <- initial_split(clickstream_model, prop = 0.8, strata = hi_clicks_avg)
train <- training(split)
test  <- testing(split)

# ── 2. Recipe ─────────────────────────────────────────────────────────────────

rec <- recipe(hi_clicks_avg ~ main_category + color + location + 
                model_photography + page + month + clothing_model + price,
              data = train) %>%
  step_lencode_glm(clothing_model, outcome = vars(hi_clicks_avg)) %>%  # target encode high-cardinality
  step_dummy(all_nominal_predictors()) %>%                              # one-hot the rest
  step_normalize(price)                                                 # scale continuous

# ── 3. Model spec ─────────────────────────────────────────────────────────────

xgb_spec <- boost_tree(
  trees = 500,
  tree_depth = tune(),
  learn_rate = tune(),
  loss_reduction = tune(),
  min_n = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# ── 4. Workflow ───────────────────────────────────────────────────────────────

wf <- workflow() %>%
  add_recipe(rec) %>%
  add_model(xgb_spec)

# ── 5. Tune ───────────────────────────────────────────────────────────────────

set.seed(42)
folds <- vfold_cv(train, v = 5, strata = hi_clicks_avg)

xgb_grid <- grid_space_filling(
  tree_depth(),
  learn_rate(),
  loss_reduction(),
  min_n(),
  size = 20
)

tune_results <- tune_grid(
  wf,
  resamples = folds,
  grid = xgb_grid,
  metrics = metric_set(roc_auc, accuracy)
)

# ── 6. Finalize & fit ─────────────────────────────────────────────────────────

best_params <- select_best(tune_results, metric = "roc_auc")

final_wf <- finalize_workflow(wf, best_params)
final_fit <- last_fit(final_wf, split)

# ── 7. Evaluate ───────────────────────────────────────────────────────────────

collect_metrics(final_fit)

final_fit %>%
  collect_predictions() %>%
  conf_mat(truth = hi_clicks_avg, estimate = .pred_class)

# Variable importance
final_fit %>%
  extract_fit_parsnip() %>%
  vip::vip(num_features = 15)
```

```{r}
# ── ROC Curve ─────────────────────────────────────────────────────────────────
final_fit %>%
  collect_predictions() %>%
  roc_curve(truth = hi_clicks_avg, .pred_1, event_level = "second") %>%
  autoplot()

# ── Confusion Matrix (visual) ──────────────────────────────────────────────────
final_fit %>%
  collect_predictions() %>%
  conf_mat(truth = hi_clicks_avg, estimate = .pred_class) %>%
  autoplot(type = "heatmap")

preds <- final_fit %>% collect_predictions()

confusionMatrix(
  factor(preds$.pred_class, levels = c(0, 1)),
  factor(preds$hi_clicks_avg, levels = c(0, 1)),
  positive = "1"
)
```
Boosted Gradient with target encoding Accuracy : 0.6906



Now I try pBART again, but with target encoding

```{r}
# ── 1. Prep ───────────────────────────────────────────────────────────────────

clickstream_model <- clickstream %>%
  mutate(
    hi_clicks_med = as.factor(hi_clicks_med),  # must be factor for classification
    across(c(main_category, color, location, model_photography, 
             page, month, clothing_model), as.factor)
  )

set.seed(42)
split <- initial_split(clickstream_model, prop = 0.7, strata = hi_clicks_med)
train <- training(split) %>% select(hi_clicks_med, main_category, color, location, 
                                     page, country, month, clothing_model)
test  <- testing(split)  %>% select(hi_clicks_med, main_category, color, location, 
                                     page, country, month)
# Build encoding recipe on training data only
encoding_rec <- recipe(hi_clicks_med ~ ., data = train) %>%
#  step_lencode_glm(clothing_model, outcome = vars(hi_clicks_med)) %>%
  step_dummy(all_nominal_predictors()) %>%
  prep(training = train)

# Apply to train and test
xtrain_encoded <- bake(encoding_rec, new_data = train) %>% 
                      select(-hi_clicks_med) %>% as.data.frame()
xtest_encoded  <- bake(encoding_rec, new_data = test) %>% 
                      select(-hi_clicks_med) %>% as.data.frame()

# Extract outcomes
ytrain <- train$hi_clicks_med
ytest  <- test$hi_clicks_med

# Fit pbart
#barfit <- pbart(xtrain_encoded, as.numeric(as.character(ytrain)),
#                x.test = xtest_encoded,
#                k = 2, ntree = 200, nskip = 1000, ndpost = 2000)

barfit <- pbart( xtrain_encoded,as.numeric(as.character(ytrain)), x.test = xtest_encoded, 
                 k = 2, ntree = 200)
```



```{r}
# Get predicted probabilities for test set
bart_probs <- barfit$prob.test.mean

# Convert to binary predictions
bart_pred <- ifelse(bart_probs > 0.5, 1, 0)

# Confusion matrix
confusionMatrix(factor(bart_pred), factor(ytest), positive = '1')


roc_bart <- roc(as.numeric(as.character(ytest)), bart_probs)
auc(roc_bart)
plot(roc_bart, main = paste0("BART ROC Curve | AUC = ", round(auc(roc_bart),3)))
```

```{r}


# Your known original variable names
original_vars <- c("main_category", "color", "location", 
                   "price", "model_photography", "page", "month")

# Extract and aggregate importance
importance_df <- data.frame(
  dummy_var  = colnames(barfit$varcount),
  importance = colMeans(barfit$varcount)
) %>%
  mutate(original_var = case_when(
    dummy_var %in% original_vars ~ dummy_var,
    TRUE ~ str_remove(dummy_var, "\\d+$")
  )) %>%
  group_by(original_var) %>%
  summarise(importance = sum(importance)) %>%
  mutate(importance_pct = importance / sum(importance) * 100) %>%
  arrange(desc(importance))

print(importance_df)

ggplot(importance_df, aes(x = reorder(original_var, importance), y = importance_pct)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "BART Variable Importance (aggregated)",
    x = NULL,
    y = "% of total splits"
  ) +
  theme_minimal()
```