---
title: "Exploratory Data Analysis"
Author: "Alex Marcek"
output: html_document
---

# Data Import
```{r, eval = FALSE}
# Force package installation
options(download.file.method = "wininet")
install.packages("ggcorrplot")
```


We will be using the cleaned data that resulted from the processing we did in `D474_DataClean.rmd`

```{r}
library(here)

clickstream <- read.csv(here("Data", "EditedData", "clickstream_dc.csv"))

head(clickstream)
str(clickstream)
colnames(clickstream)
```
Note that data has been interpreted back into a character and integer data types. We will need to convert back to date if we wish to reference it as such. (We will need this later in the document).

```{r}
clickstream$date <- as.Date(clickstream$date)

# Same with factors, this time keep sid as 'int' type
cols <- c("hi_clicks_avg", "hi_clicks_med", "color", "model_photography", "location", "price_2", "country", "main_category", "clothing_model", "page")

clickstream[cols] <- lapply(clickstream[cols], as.factor)

str(clickstream)
```

# Univariate Analysis

```{r}
library(dplyr)
library(purrr)

# Lets take a look at our categorical variables first
categorical_cols <- clickstream %>% select(where(is.factor) | where(is.character)) %>% 
  names() %>%
  setdiff("key")

# Frequency tables
categorical_cols %>%
  set_names() %>%
  map(~ clickstream %>%
        count(.data[[.x]], sort = TRUE) %>%
        mutate(pct = round(n / sum(n) * 100, 1)) %>%
        rename(level = 1))
```
Some interesting insights from this table:

(Table 1) The majority (81%) of users appear to be from country 29, which according to our documentation is Poland, followed by the Czech Republic (10.9%). It is appropriate to say that the majority of our data comes from the Central European region.

(Table 2 & 3) People in this data set appear to focus their clicks equally among the main product categories (trousers, sale, blouses, skirts) The same can be said for the top 30 clothing models. We note that there is no steep dropoff point among clothing models, though some models are clearly more popular than others. 

(Table 4) Black and blue items are the most popular among viewers of this website

(Table 5) People appear to focus on the photos located on the top of the page and there 3.4% drop in clicks between the top middle and bottom middle (3rd most popular) picture locations. 

(Table 6) En face model photography photos tended to get more clicks

(Table 7) There is not much difference in clicks between products whose price was greater than the average price of the entire category, and products whose prices were not greater. 

(Table 8) The majority of users stayed on either page 1 or 2, and did not venture out to pages 3, 4 or 5. 

(Table 9) The majority of clicks (68.3%) are classified as "low-engagement" clicks, or below the mean value for the entire data set.


```{r}
library(ggplot2)
library(tidyr)

clickstream %>%
  select(where(is.factor) | where(is.character)) %>%
  select(-c(key, clothing_model, color, country)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "level") %>%
  count(variable, level) %>%
  group_by(variable) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  ggplot(aes(x = reorder(level, n), y = pct)) + 
  geom_col(fill = "lightblue") + 
  coord_flip() + 
  facet_wrap(~ variable, scales = "free") + labs(x = NULL, y = "Percent")
```

```{r}
clickstream %>%
  select(c(clothing_model, color, country)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "level") %>%
  count(variable, level) %>%
  group_by(variable) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  ggplot(aes(x = reorder(level, n), y = pct)) + 
  geom_col(fill = "lightblue") + 
  coord_flip() + 
  facet_wrap(~ variable, scales = "free") + 
  labs(x = NULL, y = "Percent") + 
  theme(axis.text.y = element_text(size = 4))
```

Now, lets take a look at our quantitative features
```{r}
# Summary
clickstream %>% 
  select(where(is.numeric)) %>%
  summary()
```

```{r}
# Visual representation of numerical variables
clickstream %>% select(where(is.numeric)) %>% 
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value)) + 
  geom_histogram(bins = 30, fill = "lightblue") + 
  facet_wrap(~ variable, scales = "free") + 
  labs(x = NULL, y = "Count") + 
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5))
```
As we can see from our visuals, a lot of our variables seem to be exponentially distributed or uniformly distributed. The only variable that appears to have an approximately normal distribution is `price`. We may need to perform some data transformations later on. I don't have much concern over the uniformly distributed variable, since for a lot of them, it makes sense that they are uniformly distributed (like `sid`). It is important to take note now, which ones appear to be exponentially distributed, so that we know how to transform them if we encounter performance issues later on: (`hi_clicks`, `model_photography`, `page`, `clothing_model`, `color`, `country`, and `order`).

# Multivariate Analysis

```{r}
str(clickstream)
```
When considering bivariate analysis on data that has both quantitative and categorical data, we must consider multiple cases for multivariate / bivariate EDA. We will split into 3 cases: Categorical vs Categorical, Categorical vs Quantitative, Quantitative vs Quantitative. 

Setup: 
```{r}
library(tidyr)
library(purrr)

categorical_cols <- c("country", "main_category", "clothing_model", "color", "location", "model_photography", "price_2", "page", "hi_clicks_avg", "hi_clicks_med")

quantitative_cols <- c("sid", "order", "price" )
```


(1) Quantitative vs Quantitative

In this case, since we only have 3 quantitative variables, we can simply find their correlation matrix: 
```{r}
library(ggcorrplot)

cor_matrix <- clickstream %>%
  select(all_of(quantitative_cols)) %>%
  cor(use = "complete.obs")

ggcorrplot(
  cor_matrix,
  method = "circle",
  type = "lower",
  lab = TRUE,
  colors = c("red", "green", "blue"),
  title = "Feature corr matrix"
)
```
No apparent issues with correlation amongst quantitative variables.

(2) Categorical vs Categorical

We use Chi-sq and Cramer's V for all categorical pairs to quantify association across all categorical pairs.
```{r}
library(tibble)

cramers_v <- function(x, y) {
  tbl <- table(x, y)
  chi2 <- suppressWarnings(chisq.test(tbl))$statistic
  n <- sum(tbl)
  k <- min(nrow(tbl), ncol(tbl))
  sqrt(chi2 / (n * (k - 1)))
}

# All pairs
cat_pairs <- combn(categorical_cols, 2, simplify = FALSE)

assoc_results <- map_dfr(cat_pairs, function(pair) {
  tibble(
    var1 = pair[1],
    var2 = pair[2],
    cramers_v = round(cramers_v(clickstream[[pair[1]]], clickstream[[pair[2]]]), 3)
  )
}) |>
  arrange(desc(cramers_v))

print(assoc_results)

# Visualize as heatmap
assoc_matrix <- assoc_results |>
  bind_rows(rename(assoc_results, var1 = var2, var2 = var1)) |>
  bind_rows(tibble(var1 = categorical_cols, var2 = categorical_cols, cramers_v = 1)) |>
  pivot_wider(names_from = var2, values_from = cramers_v)

assoc_mat <- as.matrix(column_to_rownames(assoc_matrix, "var1"))

ggcorrplot(assoc_mat,
  lab      = TRUE,
  lab_size = 3,
  colors   = c("white", "lightblue", "blue"),
  title    = "Cramer's V — categorical associations",
) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
We see a few variables with strong to perfect association among categorical variables according to Cramer's V.

(3) Quantitative vs Categorical

The best way that we can show correlation between Quantitative vs Categorical variables is by using a Kruskal-Wallis across all categorical variables. This will show which categorical variables meaningfully stratify our quantitative variables

```{r}
kw_all <- map_dfr(categorical_cols, function(cat) {
  map_dfr(quantitative_cols, function(quant) {
    formula <- as.formula(paste(quant, "~", cat))
    test <- tryCatch(
      kruskal.test(formula, data = clickstream),
      error = function(e) NULL
    )
    if (is.null(test)) return(NULL)
    tibble(
      categorical = cat,
      quantitative = quant,
      p_value     = round(test$p.value, 4),
      significant = test$p.value < 0.05
    )
  })
})

p_value <- 0.05 # standard

# Visualise as significance heatmap
kw_all |>
  mutate(neg_log_p = -log10(p_value + 1e-10)) |>
  ggplot(aes(x = quantitative, y = categorical, fill = neg_log_p)) +
  geom_tile() +
  geom_text(aes(label = ifelse(significant, "*", "")),
            size = 5, colour = "white") +
  scale_fill_gradient(low = "lightblue", high = "blue",
                      name = "-log10(p)") +
  labs(title = "Kruskal-Wallis significance heatmap",
       subtitle = "* = significant at p < 0.05",
       x = NULL, y = NULL) +
  theme_minimal()
```
For the dark blue sections, there is strong evidence that the categorical variable in question splits quantitative variables into different distributions. So, we see strong evidence of the majority of our categorical variables meaningfully stratifying our quantitative variables. For dark rows, order, price and sid vary meaningfully across that category. For dark columns, that quantitative variables is sensitive to  many categorical groupings. For instance, price varies signficantly by each of our categorical variables. 


```{r}
par(mfrow = c(1, 2))

barplot(
  table(as.integer(format(clickstream$date, "%m"))),
  main = "Frequency of price_2",
  xlab = "Category",
  ylab = "Count",
  col = "lightblue"
)

barplot(
  table(clickstream$country),
  main = "Frequency of price_2",
  xlab = "IP",
  ylab = "Count",
  col = "lightblue"
)
as.data.frame(table(clickstream$country)) %>% arrange(desc(Freq)) %>% head(5)
```
Months have roughly the same number of interactions.
Top 5 most popular IPs go to: Poland, Czech Republic, Lithuania, NA, NA.