---
title: "RNN"
author: "Alex Marcek"
output: html_document
---

## Data Import
```{r, eval = FALSE}
# Force package installation
options(download.file.method = "wininet")
install.packages("torch")
```


```{r}
library(here)

clickstream <- read.csv(here("Data", "EditedData", "clickstream_dc.csv"))

str(clickstream)
```
Make sure we have factors and date times where we want them. 
```{r}
# Datetime
clickstream$date <- as.Date(clickstream$date)

# Factors
cols <- c("color", "model_photography", "location", "price_2", "country", "sid", "main_category", "clothing_model", "page")

clickstream[cols] <- lapply(clickstream[cols], as.factor)

str(clickstream)
# we will treat hi_clicks_avg and hi_clicks_med as integers here
```

## Data Restructuring
Preamble: we aim to classify users as high engagement or low engagement based on some feature set. In this case, we will be engineering some features
```{r}
# Data preparation
library(dplyr)
library(purrr)
library(tidyr)

# For each session, we want to summarize clicks, main_category, clothing_models, pages and item colors.
session_features <- clickstream %>%
  group_by(sid) %>%
  summarize(
    # Tallies
    n_clicks = n(),
    n_categories = n_distinct(main_category),
    n_items = n_distinct(clothing_model),
    n_pages = n_distinct(page),
    n_colors = n_distinct(color),
    n_locations = n_distinct(location),
    
    # Price related behavior
    mean_price = mean(price),
    sd_price = sd(price), # May encounter NA's here by construction (1-click case)
    price_range = max(price) - min(price),
    min_price = min(price),
    max_price = max(price),
    
    # Engagement
    hi_clicks_avg = mean(hi_clicks_avg), # interpret as pct of sessions clicks that classified as high using mean click classifier
    hi_clicks_med = mean(hi_clicks_med), # interpret as pct of sessions clicks that classified as high using median click classifier
    
    # User behavior
    revisit_rate = mean(duplicated(clothing_model)), # useful for looping website navigation behavior
    category_switches = sum(main_category != lag(main_category, default = first(main_category))), # Checks to see if users switch main category between clicks, tallies switches
    .groups = "drop"
  )

head(session_features)
```
Note that we have missing data in the sd_price column. We will summarize here: 

```{r}
# lets check for percentage of missing data in sd_price
sum(as.integer(is.na(session_features$sd_price))) / nrow(session_features)
```
Around 20 percent of our data!
```{r}
session_features %>%
  filter(is.na(sd_price)) %>%
  summarize(
    n = n(),
    all_single = all(n_clicks == 1),
    max_clicks = max(n_clicks)
  )
```
We see that in all cases where sd_price is NA, it always results in a single click session. In this case, we know that it is appropriate to inpute 0 for the NAs in sd_price. This is because the NAs resulted as a computational artifact from the `sd()` function, and a true representation of the sd's for single click session is truly zero, not undefined. Lets inpute: 
```{r}
session_features <- session_features %>%
  mutate(sd_price = replace_na(sd_price, 0))
```

We may want to consider separating these single click session ids from focused users later on in the process, since the number of single click users is ~20%. We will consider this when further classifying our users. For now, we will maintain that single click users are indeed "low engagement"

Now we will create our binary label. 
```{r}
median_clicks <- median(session_features$n_clicks) # use median to avoid (dist) assumptions associated with mean value threshold

session_features <- session_features %>%
  mutate(engagement = ifelse(n_clicks >= median_clicks , "high", "low"))

table(session_features$engagement)
```
Observe we now have a 54.29% , 45.71% split now. This opposes our previous 68.3% , 31.7% split of low engagement vs high engagement "clicks". The difference is now we are looking directly at session id. This aims to address the majority-class issue that we encountered early on in our model fitting procedures 

```{r}
library(ggplot2)

ggplot(session_features, aes(n_clicks, fill = engagement)) + 
  geom_histogram(bins = 40, alpha = 0.8) + 
  geom_vline(xintercept = median_clicks, linetype = "dashed") + 
  scale_fill_manual(values = c("high" = "lightgreen", "low" = "lightblue")) + 
  labs(title = "Click Distribution by Session", 
       x = "Number of Clicks", y = "Sesssions")
```
If we are considering a simple unsupervised approach, this median split rule is enough to classify users as either high or low engagement. This is confirmed in part by the previous models that we fit. They only performed slightly better than this simple classification rule, when attempting to classify users as "high" or "low" engagement. 

A more interesting question is a more detailed classification of our user base. The previous models that we have build have focused on classifying users as high or low engagement (number of clicks in a session). We will look at two separate interesting cases that tell us a little more about the users, and how they may interact with the website. First, we will further classify users as focused, explorers and price sensitive users. Later on, we will also look at predicting the users next click based on their previous clicks. This is useful information for companies, especially for targeted advertising. 

Lets expand on what it means for a user to be classified as focused, explorer or price sensitive. We say a user is (focused) if their number of clicks is low, and they stay in consistent categories. We say a user is an (explorer) if their click count is high, and they visit many diverse categories. Finally, we call a user price sensitive if they stay in consistently high, or low priced sections. In this section, using knn, we will focus on the classification question: focused, explorer or price sensitive.

(1) First, scale data
```{r}
# Define which variable we want to consider
cluster_vars <- c("n_clicks", "n_categories", "n_items", "n_pages", "mean_price", "sd_price", "price_range", "hi_clicks_med", "revisit_rate", "category_switches")

X <- session_features %>%
  select(all_of(cluster_vars)) %>%
  scale()

scale_center <- attr(X, "scaled:center")
scale_sd <- attr(X, "scaled:scale")
```

Since we are in an unsupervised learning case, it is more difficult to implement cross validation. We have no assigned response, we are treating "hi_clicks_med" as another predictor. Bootstrapping can be implemented, but there is a high computational cost. Some recommended methods for choosing k in a knn, would be either elbow, or silhouette. 

(2) Then choose optimal k for our k-Nearest-Neighbors
```{r}
library(factoextra)
library(cluster)

set.seed(050926)

# Elbow method
fviz_nbclust(X, kmeans, method = "wss") + 
  labs(title = "Elbow Method")

# Silhouette
fviz_nbclust(X, kmeans, method = "silhouette") + 
  labs(title = "Silhouette Method")
```
Our optimal k appears to be 2 according to the silhouette method, and either 2 or 3 for the elbow method. We will choose 2, and evaluate performance. 

(3) Fit k-means
```{r}
set.seed(050926)

km <- kmeans(X, centers = 2, nstart = 50, iter.max = 300)

session_features <- session_features %>%
  mutate(cluster = factor(km$cluster)) # create cluster labels coding
```

(4) Label clusters
```{r}
cluster_profiles <- session_features %>%
  group_by(cluster) %>%
  summarize(
    n = n(),
    avg_clicks = mean(n_clicks),
    avg_cats = mean(n_categories),
    avg_items = mean(n_items),
    avg_price = mean(mean_price),
    avg_price_var = mean(sd_price), 
    avg_revisit = mean(revisit_rate),
    avg_hi_clicks = mean(hi_clicks_avg),
    avg_switches = mean(category_switches),
    .groups = "drop"
  )

print(cluster_profiles)
```
```{r}
# Create new user classification labels
cluster_labels <- c("1" = "Explorer",
                    "2" = "Focused browser",
                    "3" = "Price sensitive")

session_features <- session_features %>%
  mutate(segment = recode(cluster, !!!cluster_labels))
```

(5) Validate
```{r}
sil <- silhouette(km$cluster, dist(X))
cat("Mean silhouette:", round(mean(sil[, 2]), 2), "\n")
fviz_silhouette(sil) + theme_minimal()

# PCA projection
pca <- prcomp(X, center = FALSE, scale. = FALSE)
pcs <- as.data.frame(pca$x[ , 1:2])
pcs$segment <- session_features$segment
```

```{r}
ggplot(pcs, aes(x = PC1, y = PC2, color = segment)) + 
  geom_point(alpha = 0.35, size = 1.2) + 
  stat_ellipse(level = 0.85, linewidth = 1) + 
  scale_color_manual(values = c(
    "Explorer" = "lightgreen",
    "Focused browser" = "lightblue",
    "Price sensitive" = "red"
  )) +
  labs(title = "Session segments - PCA projection") + 
  theme_minimal()
```
After running this KNN silhouette and pca projection evaluation, we see that we are still in the high vs low click classification problem. What could be happening is k = 2 may be separating singular clicks (just detecting bounces), and not actually telling us anything about browsing style. Lets check this: 

```{r}
km2 <- kmeans(X, centers = 2, nstart = 50)
session_features$cluster2 <- factor(km2$cluster)

session_features %>%
  group_by(cluster2) %>%
  summarise(
    n = n(),
    avg_clicks = mean(n_clicks),
    avg_cats = mean(n_categories),
    avg_price_var = mean(sd_price),
    .groups = "drop"
  )
```
Since there is such a large difference in average clicks between cluster 1 and cluster 2, this is why silhouette is preferring k = 2. Lets check how many of our clicks are exactly 1, and how many are below 3.

```{r}
session_features %>%
  group_by(cluster2) %>%
  summarize(
    pct_single_click = mean(n_clicks == 1),
    pct_under_3 = mean(n_clicks <= 3),
    median_clicks = median(n_clicks),
    .groups = "drop"
  )
```
There's a clear problem here! Close to 30% of our cluster 2, or "focused" users are single clicks, and 64 percent of them are under 3 clicks! Lets filter to only include users that are above three clicks. This way we are including users that are at least minimally engaged in the website (define as greater than 3 clicks). 

```{r}
X_mult <- X[session_features$n_clicks > 3, ]

set.seed(051026)

fviz_nbclust(X_mult, kmeans, method = "silhouette", k.max = 8) + 
  labs(title = "Silhouette - greater than 3 clicks") + theme_minimal()
```
Still optimal value of k = 2, lets try a slightly higher cutoff and see what happens

```{r}
X_mult <- X[session_features$n_clicks > 9, ]

set.seed(051026)

fviz_nbclust(X_mult, kmeans, method = "silhouette", k.max = 8) + 
  labs(title = "Silhouette - greater than 9 clicks") + theme_minimal()
```
This provides some valuable insight. It appears that regardless of the cutoff we choose (9 being maximal since it is the avg cutoff that we use in the naive approach), our data may be genuinely bimodal. Lets check one last thing to be sure. Our engineered features may be strongly correlated. Lets see if we can diagnose which (if any) are.

```{r}
cor_clicks <- cor(session_features[, cluster_vars],
                  use = "complete.obs")[, "n_clicks"]

sort(cor_clicks, decreasing = TRUE)
```
As we see, `n_items`, `hi_clicks_med`,  `category_switches`, `price_range`, `n_pages`, `n_categories` and `sd_price` are all highly correlated with `n_clicks`. Lets try to normalize to decorrelate

```{r}
session_features_norm <- session_features %>%
  mutate(
    cats_per_click = n_categories / n_clicks,
    items_per_click = n_items / n_clicks,
    pages_per_click = n_pages / n_clicks,
    switches_per_click = category_switches / n_clicks,
    price_var_per_click = sd_price / (n_clicks + 1)
  )

style_vars <- c("cats_per_click", "items_per_click", "pages_per_click", "switches_per_click", "price_var_per_click", "revisit_rate", "mean_price")

# Check again
cor_norm <- session_features_norm %>%
  filter(n_clicks > 3) %>%
  select(n_clicks, any_of(style_vars)) %>%
  cor(use = "complete.obs")

round(cor_norm[, "n_clicks"], 3)
```
This has decorrelated pretty well, now lets check the silhouette again.

```{r}
X_style <- session_features_norm %>%
  filter(n_clicks > 3) %>%
  select(all_of(style_vars)) %>%
  scale()

set.seed(051026)
fviz_nbclust(X_style, kmeans, method = "silhouette", k.max = 8) + 
  labs(title = "Silhouette - rate-normalized features") + 
  theme_minimal()
```
Finally, we can facilitate the implementation of more than 2 classifications. We proceed: 

```{r}
set.seed(051026)

km_style <- kmeans(X_style, centers = 3, nstart = 50, iter.max = 300)

session_features_norm <- session_features_norm %>%
  filter(n_clicks > 3) %>%
  mutate(style_cluster = factor(km_style$cluster))
```

```{r}
# Cluster profiling
cluster_profiles <- session_features_norm %>%
  group_by(style_cluster) %>%
  summarise(
    n = n(),
    avg_cats_per_click = mean(cats_per_click),
    avg_items_per_click = mean(items_per_click), 
    avg_pages_per_click = mean(pages_per_click), 
    avg_switches = mean(switches_per_click), 
    avg_price_var = mean(price_var_per_click), 
    avg_revisit = mean(revisit_rate), 
    avg_price_tier = mean(mean_price),
    avg_clicks = mean(n_clicks),
    avg_categories = mean(n_categories),
    .groups = "drop"
  )

print(cluster_profiles)
```
After reading the above output we can classify our clusters, recall we want explorers to have high: avg_cats_per_click, avg_switches; low: avg_revisit. Focused browsers to have high: avg_revisit, avg_items_per_click; low: avg_switches, avg_cats_per_clicks. And we want price sensitive to have high: avg_price_var, avg_price_tier; low: avg_switches.

```{r}
cluster_labels <- c(
  "1" = "Price sensitive",
  "2" = "Focused browser",
  "3" = "Explorer"
)

session_features_norm <- session_features_norm %>%
  mutate(segment = recode(style_cluster, !!!cluster_labels))
```

Now we validate our above model with silhouette:
```{r}
sil <- silhouette(km_style$cluster, dist(X_style))
cat("Mean silhouette width:", round(mean(sil[, 3]), 3), "\n")

fviz_silhouette(sil) +
  labs(title = "Silhouette — k=3 style clusters") +
  theme_minimal()
```
Since our mean silhouette width achieves a value of 0.259, we say that it has reasonable structure. 

Lets validate with bootstrap stability
```{r}
library(fpc)

set.seed(051026)

stability <- clusterboot(
  X_style,
  B = 100,
  clustermethod = kmeansCBI,
  k = 3,
  seed = 051026
)

cat("cluster stability: \n")
print(round(stability$bootmean, 3))
```
We achieved high cluster stability for all three of our clusters, meaning they are each a genuine behavioral group. 

```{r}
# PCA projection
pca <- prcomp(X_style, center = FALSE, scale. = FALSE)
pcs <- as.data.frame(pca$x[, 1:2])
pcs$segment <- session_features_norm$segment

ggplot(pcs, aes(x = PC1, y = PC2, colour = segment)) +
  geom_point(alpha = 0.35, size = 1.2) +
  stat_ellipse(level = 0.85, linewidth = 1) +
  scale_colour_manual(values = c(
    "Explorer"         = "blue",
    "Focused browser"  = "green",
    "Price-sensitive"  = "red"
  )) +
  labs(title = "Browsing style segments — PCA projection") +
  theme_minimal()

# Radar chart of cluster profiles
library(fmsb)

radar_data <- cluster_profiles |>
  select(avg_cats_per_click, avg_items_per_click,
         avg_switches, avg_price_var,
         avg_revisit, avg_price_tier) |>
  as.data.frame()

rownames(radar_data) <- cluster_labels

radar_data <- rbind(
  apply(radar_data, 2, max),
  apply(radar_data, 2, min),
  radar_data
)

radarchart(
  radar_data,
  pcol  = c("blue", "green", "red"),
  pfcol = adjustcolor(c("blue", "green", "red"), 0.2),
  plwd  = 2,
  cglcol = "grey",
  title = "Behavioral profiles per segment"
)
legend("topright",
  legend = cluster_labels,
  col    = c("blue", "green", "red"),
  lty = 1, lwd = 2, bty = "n"
)
```

Now that we have classified 3 clusters, we must now evaluate the clusters performance. For our clustering to be good, we must have cohesion, separation and stability. 

(1) Measuring silhouette width tells us about cohesion and separation. 

```{r}
library(cluster)

sil <- silhouette(km_style$cluster, dist(X_style))

# Overall
cat("Mean silhouette width:", round(mean(sil[, 3]), 3), "\n")

# Per - cluster
cluster_sil <- aggregate(sil[, 3], by = list(cluster = sil[, 1]), mean)
print(cluster_sil)

# Full plot
fviz_silhouette(sil) + 
  labs(title = "Sil width per session") + theme_minimal()
```
As we can see from our plot, our clusters reach close to the same peak, however clusters 2 and 3 have some negative values, This suggests that these (negative valued) session are closer to a neighboring cluster than their own. So, these are likely misclassifications. We also have a mean width of 0.259, indicating reasonable structure of our clusters, but with some overlap. 

Next, we will check cohesion using within-cluster sum of squares and WSS
```{r}
cat("Total WSS:", round(km_style$tot.withinss, 2), "\n")
cat("Per-clust WSS:\n")
print(round(km_style$withinss, 2))

bss_ratio <- km_style$betweenss / km_style$totss
cat("Between sum of squares ratio:", round(bss_ratio , 3), "\n")
```
Our model achieved a between cluster sum of squares ratio of 0.407. A reasonable value is often considered to be above 0.6. So, our cluster structure falls short of being reasonably cohesive. The interpretation of between-cluster sum of squares, intuitively, is the proportion of total variance explained by our cluster structure. 

Finally, above, we already tested our Bootstrap stability. The model achieved a score of 0.987, 0.984, 0.989 for clusters 1, 2 and 3 respectively. This means that our clusters are highly stable, and are real clusters, not just a k-means artifact. So, we know that our clusters are real and consistent, but due to the structure of our \texbtf{data itself}, are not very distinct. This points us back to the fact that user interaction with the website is able to be classified simply using session clicks counts, and the two distinct classes are truly "high" and "low" engagement users classified using the rule "low" is below median, "high" is above median click counts across all users. 

## Model Evaluations: New Data Structure
Now that we have collapsed and restructured our data, as well as taking care of some correlation issues (in first modification iteration), lets revisit some of our older models and evaluate their performance on the updated data.

```{r}
# Data partitioning
set.seed(051026)
n <- nrow(session_features_norm)
train_idx <- sample(n, floor(0.8 * n))
test_idx  <- setdiff(seq_len(n), train_idx)

train_data <- session_features_norm[train_idx, ]
test_data  <- session_features_norm[test_idx, ]
```

Multinomial Logistic Regression (New linear baseline) 
```{r}
library(nnet)

model_logit <- multinom(segment~ 
                          cats_per_click + items_per_click + switches_per_click +
                          price_var_per_click + revisit_rate + mean_price,
  data  = train_data,
  trace = FALSE
)

pred_logit <- predict(model_logit, test_data)
```

Linear Discriminant Analysis
```{r}
library(MASS)

model_lda <- lda(segment~ 
                   cats_per_click + items_per_click + switches_per_click + 
                   price_var_per_click + revisit_rate + mean_price,
  data = train_data
)

pred_lda <- predict(model_lda, test_data)$class
```

Random Forests
```{r}
library(randomForest)

model_rf <- randomForest(factor(segment)~ 
                           cats_per_click + items_per_click + switches_per_click + 
                           price_var_per_click + revisit_rate + mean_price,
  data   = train_data,
  ntree  = 500
)

pred_rf <- predict(model_rf, test_data)
```

Comparisons in context of the classification problem above, on same metrics
```{r}
library(caret)

actual <- test_data$segment

results <- list(
  logistic = confusionMatrix(factor(pred_logit), factor(actual)),
  lda      = confusionMatrix(factor(pred_lda), factor(actual)),
  rf       = confusionMatrix(factor(pred_rf), factor(actual))
)

# Extract key metrics into a comparison table
comparison <- map_dfr(names(results), function(model_name) {
  cm <- results[[model_name]]
  tibble(
    model = model_name,
    accuracy = cm$overall["Accuracy"],
    kappa = cm$overall["Kappa"],
    macro_f1 = mean(cm$byClass[, "F1"], na.rm = TRUE)
  )
})

print(comparison)
```
Note that we do not use our usual metrics like training or test MSE. This is intentional, because we are still interpreting the relative performance of our models in an unsupervised case. At this point, we have not assigned class labels. The interpretation of our results are, since Random forest has approximately the same performance as LDA, the segmented boundaries are linear, and the k-means procedure that we detailed above had genuine linear structure. Also, since all of the models performed well, we have strong cluster structure and our results will be reproducible!

## Exension of Study: RNN fit
In this section we focus on a different question that stem from our initial study. Another useful tool that the website can use is predicting the next click of users based on their previous browsing session behavior. This is useful for targeted advertising and website navigation optimization that can be utilized by frontend developers. 

We can use the clustering results we found above to enrich our RNN.
Here's what the RNN with clustering sees: 

[click_1, click_2, ..., click_N] + segment_label -> click_N+1 prediction

Recurrent Neural Network (RNN): Data preparation
```{r}
# Data preparation
cs_model_ds <- clickstream %>%
  left_join(
    session_features_norm %>%
      select(sid, segment),
    by = "sid"
  ) %>%
  filter(!is.na(segment)) %>% # excluding bounce sessions 
  arrange(sid, order) 

cs_model_ds <- cs_model_ds %>%
  mutate(segment_id = as.integer(factor(segment)))

segment_vocab_size <- n_distinct(cs_model_ds$segment_id)

## Now we have a data set (cs_model_ds) that can be used in the RNN context
```

RNN: Feature encoding
```{r}
categorical_cols <- c("main_category", "country", "clothing_model", "color", "location", "model_photography", "page")

encoders <- list()
for (col in categorical_cols) {
  lvls <- unique(cs_model_ds[[col]])
  encoders[[col]] <- lvls
  cs_model_ds[[col]] <- as.integer(factor(cs_model_ds[[col]], levels = lvls))
}

# Set threshold for rare categories
min_freq <- 5
for (col in categorical_cols){
  freq <- table(cs_model_ds[[col]])
  rare <- as.integer(names(freq[freq < min_freq]))
  cs_model_ds[[col]][cs_model_ds[[col]] %in% rare] <- 0L
}

# Encoding for numeric features
num_cols <- c("price", "hi_clicks_avg", "hi_clicks_med")
for (col in num_cols){
  cs_model_ds[[col]] <- scale(cs_model_ds[[col]])[, 1]
}

# Engineer cyclical date features
library(lubridate)
cs_model_ds <- cs_model_ds %>%
  mutate(
    dow_sin = sin(2 * pi * wday(date) / 7), # day of week
    dow_cos = cos(2 * pi * wday(date) / 7),
    month_sin = sin(2 * pi * month(date) / 12),
    month_cos = cos(2 * pi * month(date) / 12),
  ) %>%
  select(-date)

all_num_cols <- c(num_cols, "dow_sin", "dow_cos", "month_sin", "month_cos")
```

RNN: Sequencing
```{r}
library(purrr)

sessions <- cs_model_ds %>%
  group_by(sid) %>%
  group_split()

make_seqs <- function(session){
  n <- nrow(session)
  if (n < 2) return(NULL)
  
  map(2:n, function(i){
    list(
      input = session[1:(i - 1), ], # this is our i-1 pathway
      target = session$main_category[i],
      segment_id = session$segment_id[1]
    )
  })
} 
```

```{r}
all_seqs <- sessions %>%
  map(make_seqs) %>%
  compact() %>%
  list_flatten()
```

RNN: Split and Train
```{r}
set.seed(051026)

all_sids <- unique(cs_model_ds$sid)
n <- length(all_sids)
train_sids <- sample(all_sids, floor(0.7 * n))
remaining <- setdiff(all_sids, train_sids)
val_sids <- sample(remaining, floor(0.5 * length(remaining)))
test_sids <- setdiff(remaining, val_sids)

get_sid <- function(seq) unique(seq$input$sid)[1]

train_seq <- keep(all_seqs, ~get_sid(.x) %in% train_sids)
val_seq   <- keep(all_seqs, ~get_sid(.x) %in% val_sids)
test_seq  <- keep(all_seqs, ~get_sid(.x) %in% test_sids)
```

RNN: Dataset and Loader
```{r}
library(torch)

MAX_LEN <- 20
NUM_NUMERIC  <- length(all_num_cols)

click_dataset <- dataset(
  name = "ClickDataset",

  initialize = function(sequences) {
    self$sequences <- sequences
  },

  .getitem = function(i) {
    seq   <- self$sequences[[i]]
    inp   <- seq$input
    n     <- nrow(inp)

    # Truncate 
    if (n > MAX_LEN) inp <- tail(inp, MAX_LEN)
    pad   <- MAX_LEN - nrow(inp)

    # Numeric tensor 
    num_mat <- as.matrix(inp[, all_num_cols])
    if (pad > 0) # padding
      num_mat <- rbind(matrix(0, pad, ncol(num_mat)), num_mat)

    # Categorical tensors
    cat_tensors <- lapply(categorical_cols, function(col) {
      vals <- inp[[col]]
      if (pad > 0) vals <- c(rep(0L, pad), vals) # padding
      torch_tensor(vals, dtype = torch_long())
    })

    list(
      cats      = categorical_tensors,
      num       = torch_tensor(num_mat, dtype = torch_float()),
      segment   = torch_tensor(seq$segment_id, dtype = torch_long()),
      target    = torch_tensor(seq$target - 1L, dtype = torch_long())
    )
  },

  .length = function() length(self$sequences)
)

train_ds <- click_dataset(train_seq)
val_ds   <- click_dataset(val_seq)
test_ds  <- click_dataset(test_seq)

train_dl <- dataloader(train_ds, batch_size = 64, shuffle = TRUE)
val_dl   <- dataloader(val_ds,   batch_size = 128)
test_dl  <- dataloader(test_ds,  batch_size = 128)
```

RNN: Model def

- Note: below three chunks are index fixes. We switch 1 to <UNK> token
```{r}
for (col in categorical_cols) {
  lvls <- unique(cs_model_ds[[col]])
  encoders[[col]] <- lvls
  # Encode starting at 2, reserve 1 for UNK
  cs_model_ds[[col]] <- as.integer(factor(cs_model_ds[[col]],
                                           levels = lvls)) + 1L
}

min_freq <- 5
for (col in categorical_cols) {
  freq <- table(cs_model_ds[[col]])
  rare <- as.integer(names(freq[freq < min_freq]))
  cs_model_ds[[col]][cs_model_ds[[col]] %in% rare] <- 1L
}

# Padding in sequences also becomes 1
# Update vocab sizes — now max index is n_levels + 1
vocab_sizes <- sapply(categorical_cols, function(col)
  max(cs_model_ds[[col]], na.rm = TRUE))
```

```{r}

```

```{r}
click_rnn <- nn_module(
  "ClickRNN",

  initialize = function(vocab_sizes,
                        embed_dim       = 16,
                        num_numeric     = NUM_NUMERIC,
                        hidden_dim      = 128,
                        num_classes,
                        segment_vocab   = 3,
                        segment_embed   = 8,
                        dropout         = 0.3) {

    self$embeddings <- nn_module_list(
      lapply(vocab_sizes, function(v)
        nn_embedding(v + 1L, embed_dim))
    )

    self$segment_embed <- nn_embedding(segment_vocab + 1L, segment_embed)

    lstm_input_dim <- length(vocab_sizes) * embed_dim + num_numeric

    self$lstm    <- nn_lstm(lstm_input_dim, hidden_dim,
                            batch_first = TRUE, dropout = dropout,
                            num_layers  = 2)
    self$dropout <- nn_dropout(dropout)

    #LSTM hidden + segment embedding
    self$classifier <- nn_linear(hidden_dim + segment_embed, num_classes)
  },

  forward = function(cats, num, segment) {
    # Embed each categorical click feature
    embeds <- lapply(seq_along(cats), function(i)
      self$embeddings[[i]](cats[[i]]) 
    )

    # Concatenate click features along feature dimension
    x <- torch_cat(c(embeds, list(num)), dim = -1)

    # Run through LSTM, take last timestep
    out <- self$lstm(x)
    h <- out[[1]][ , dim(out[[1]])[2], ]   # (batch, hidden_dim)
    h <- self$dropout(h)

    # Embed session-level segment and concatenate
    seg <- self$segment_embed(segment) # (batch, segment_embed)
    h <- torch_cat(list(h, seg), dim = -1) # (batch, hidden + seg_embed)

    self$classifier(h) # (batch, num_classes)
  }
)
```

RNN: Training loop
```{r, eval = FALSE}
vocab_sizes <- sapply(categorical_cols, function(col)
  max(cs_model_ds[[col]], na.rm = TRUE))

num_classes <- max(cs_model_ds$main_category, na.rm = TRUE)

model <- click_rnn(
  vocab_sizes = vocab_sizes,
  embed_dim = 16,
  hidden_dim = 128,
  num_classes = num_classes,
  segment_vocab = segment_vocab_size,
  segment_embed = 8,
  dropout = 0.3
)

optimizer <- optim_adam(model$parameters, lr = 1e-3)
loss_fn <- nn_cross_entropy_loss()
scheduler <- lr_step(optimizer, step_size = 5, gamma = 0.5)

best_val_loss <- Inf

for (epoch in 1:30) {

  # Training
  model$train()
  train_loss <- 0

  coro::loop(for (batch in train_dl) {
    optimizer$zero_grad()
    logits <- model(batch$cats, batch$num, batch$segment)
    loss <- loss_fn(logits, batch$target)
    loss$backward()
    nn_utils_clip_grad_norm_(model$parameters, max_norm = 1.0)
    optimizer$step()
    train_loss <- train_loss + loss$item()
  })

  # Validation
  model$eval()
  val_loss <- 0

  with_no_grad({
    coro::loop(for (batch in val_dl) {
      logits <- model(batch$cats, batch$num, batch$segment)
      val_loss <- val_loss + loss_fn(logits, batch$target)$item()
    })
  })

  scheduler$step()

  cat(sprintf("Epoch %2d | Train loss: %.4f | Val loss: %.4f\n",
              epoch, train_loss, val_loss))

  # Save best model
  if (val_loss < best_val_loss) {
    best_val_loss <- val_loss
    torch_save(model, "best_click_rnn.pt")
  }
}
```


## Future Study : Interpretability vs Performance

Some things that we consider for extending our study: troubleshooting and further defining Recurrent Neural Network procedures will allow us to ask more sophisticated questions about user interactions. However, it will fundamentally pull us out of the unsupervised case, since we will need to be defining a target variable. The way our data is constructed suggests that there is not much of a case for supervised methods without feature engineering. This is also the case for LDR, Logistic and RF. The way our data itself is constructed does not have checkout results, or specifications on its users, making target parameters not exceedingly apparent. The data that we have used so far is technically self-supervised, and in the case that we use RNN to predict next click, this is also self supervised. We can also extend our study by finding another data set with a pre-defined response variables and putting it though the same pipeline. 

We can also extend this study's interpretability by approaching the next-click problem from a stochastic process viewpoint. We would look at this through the lens of a markov chain (discrete finite state space, discrete time) and finding its probability transitions matrix. This will allow us to comment on the individual probabilities to transition between states (pages, clothing categories, etc), as well as long term, equilibrium distribution behavior. It may also be able to frame this as a markov jump process problem, with website traffic, wait times, or possibly Brownian motion. All of these would help with interpretability, that the RNN would not be able to provide. We may also look into a Bayesian approach using a Directed Acyclic Graph (DAG), giving us visual interpretability.  